---
title: "Lab 06 - Functions and iterations"
author: "Fola Remi-Shodipo"
date: "`r Sys.Date()`"
output: html_document
---


### Packages and Data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
```

```{r diamon_rename}
diamonds <- diamonds %>% rename(
    depth_pct = depth,
    length_mm = x,
    width_mm = y,
    depth_mm = z,
    table_pct = table
  )
```


## Exercise 1

```{r}
# Do some preliminary investigations & cleaning
diamonds %>%
  summarise(across(c(carat:depth_mm), mean, na.rm = TRUE))

# 1. Length vs Depth (Scatter plot)
ggplot(diamonds, aes(x = length_mm, y = depth_mm)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "Length vs Depth", x = "Length (mm)", y = "Depth (mm)")

# 2. Length vs Width (Scatter plot)
ggplot(diamonds, aes(x = length_mm, y = width_mm)) +
  geom_point(alpha = 0.5, color = "green") +
  labs(title = "Length vs Width", x = "Length (mm)", y = "Width (mm)")

# 3. Carat vs Price (Line graph)
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_line(color = "purple") +
  labs(title = "Carat vs Price", x = "Carat", y = "Price")

# 4. Carat and Cut (Bar graph)
ggplot(diamonds, aes(x = cut, y = carat)) +
  geom_bar(stat = "summary", fun = "mean", fill = "orange") +
  labs(title = "Average Carat by Cut", x = "Cut", y = "Average Carat")

# 5. Clarity and Carat (Bar graph)
ggplot(diamonds, aes(x = clarity, y = carat)) +
  geom_bar(stat = "summary", fun = "mean", fill = "pink") +
  labs(title = "Average Carat by Clarity", x = "Clarity", y = "Average Carat")

# 6. Summary of Average Price by Color
avg_price_by_color <- diamonds %>%
  group_by(color) %>%
  summarise(avg_price = mean(price, na.rm = TRUE))

# Summary of Average Price by Cut
avg_price_by_cut <- diamonds %>%
  group_by(cut) %>%
  summarise(avg_price = mean(price, na.rm = TRUE))

# Display summaries
print(avg_price_by_color)
print(avg_price_by_cut)

```


## Exercise 2

```{r}

diamonds <- diamonds %>%
  mutate(table_mm = table_pct * width_mm / 100)

# Overall table_mm arithmetic mean?
diamonds %>%
  summarise(mean_table_mm = mean(table_mm, na.rm = TRUE))





# table_mm arithmetic mean within each clarity category
diamonds %>%
  group_by(clarity) %>%
  summarise(mean_table_mm = mean(table_mm, na.rm = TRUE))






```

Lower clarity grades (like I1 and SI2) tend to have larger average table lengths, with I1 having the highest average table length of approximately 3.91 mm.
Higher clarity grades (like VVS1 and IF) have smaller average table lengths, with IF having one of the smallest averages at ar

```{r}
# calculate the arithmetic mean across all numeric variables

diamonds_summary <- diamonds %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE, .names = "avg_{.col}"))

diamonds_summary


```

## Exercise 3

```{r}
# Write gmean() command
gmean <- function(x) {
  # Ensure the input is numeric
  if (!is.numeric(x) || length(x) == 0) {
    stop("Input must be a numeric vector with positive values.")
  }
  
  # Check if there are any non-positive values
  if (any(x <= 0)) {
    stop("All values must be positive to compute the geometric mean.")
  }
  
  # Remove any NA values in x
  x <- na.omit(x)
  
  # Calculate the geometric mean using the logarithmic formula
  geometric_mean <- exp(mean(log(x)))
  
  return(geometric_mean)
}


```

## Exercise 4

```{r, error = TRUE}

test_data <- list(
  test1 = c(2.1, 3.8, 4.2),
  test2 = c(1, 10, 100, 1000),
  test3 = c(0, 1, 4),
  test4 = c(0.38,  0.94, -1.56),
  test5 = c(TRUE, TRUE),
  test6 = c("6", "7", "8")
)


# Create for loop to test gmean() on the above examples

for(i in seq_along(test_data)){
  cat("Testing gmean on test_data[[", i, "]]:", sep = "")
  tryCatch({
    result <- gmean(test_data[[i]])
    print(result)
  }, error = function(e) {
    print(e$message)
  })
}

```


## Exercise 5

```{r}
# Copy gmean() from Ex 3 and edit it to check the input





# Copy your code from Ex 4 to test your new command





```

## Exercise 6

```{r}
# replicate the for loop from Ex 4 using the map (or a similar) function

```


## Exercise 7

```{r}
# Overall table_mm arithmetic mean, median, and geometric mean?






# Arithmetic mean, median, and geometric mean of table_mm within each clarity category






```

_Add your comment here_


## Exercise 8 (optional)

```{r}
# Create gmean2() that computes the geometric mean 
#   that uses the original definition






```


## Exercise 9 (optional)

```{r}
# Create a function that computes the harmonic mean







```

## Exercise 10 (optional)

_Do derivation on paper, unless you know how to write LaTeX code_
